% -*- mode: Noweb; noweb-code-mode: latex-mode -*-%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% QPA Project User Documentation
% DESCRIPTION groebner.tex
%
% Copyright, 1998, 1999 Virginia Polytechnic Institute and State University.
% Copyright, 1998, 1999 Virginia Tech Path Algebra Project. All rights reserved.
%
% This file may be distributed in accordance with the stipulations existing
% in the LICENSE file accompanying this software.
%
% $Id: groebner.tex,v 1.1 2010/05/07 13:16:24 sunnyquiver Exp $
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\Chapter{Groebner Basis}

This chapter contains the declarations and implementations needed for
Gr\accent127obner basis. Currently, we do not provide algorithms to
actually compute Gr\accent127obner basis;  instead, the declarations and
implementations are provided here for GAP objects and the actual
elements of Gr\accent127obner basis are expected to be computed by external
packages such as `Opal' and `Groebner'.

\>InfoGroebnerBasis V

is the info class for functions dealing with Gr\accent127obner basis.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\Section{Constructing a Groebner Basis}

\>GroebnerBasis( <A>, <rels> ) F

The `GroebnerBasis' global function takes an algebra $A$ and 
a list of relations $rels \in A$ and creates an object in the
`IsGroebnerBasis' category. There are absolutely no checks for 
correctness in this function. Giving a set of relations 
that does not form a Gr\accent127obner basis may result in incorrect 
answers or unexpected errors.  This function is intended to be 
used by packages providing access to external Gr\accent127obner basis programs.




%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\Section{Categories and Properties of Groebner Basis}

\>IsGroebnerBasis( <object> ) P

is true when <object> is a Gr\accent127obner basis and 
`false' otherwise.

\beginexample 
\endexample


\>IsTipReducedGroebnerBasis( <GB> ) P

returns `true' if the Gr\accent127obner basis is tip reduced and 
`false' otherwise.

\beginexample 
\endexample


\>IsCompletelyReducedGroebnerBasis( <GB> ) P

returns `true' if the Gr\accent127obner basis is completely reduced and 
`false' otherwise.

\beginexample 
\endexample


\>IsHomogenousGroebnerBasis( <GB> ) P

returns `true' if the Gr\accent127obner basis is homogenous and 
`false' otherwise.

\beginexample 
\endexample


\>IsCompleteGroebnerBasis( <GB> ) P

returns `true' if the Gr\accent127obner basis is complete or 
`false' if it is not. While philosophically something that 
isn\pif t a complete Gr\accent127obner basis isn\pif t a Gr\accent127obner basis at all, 
this property can be used in conjuction with other properties 
to see if the the Gr\accent127obner basis contains enough information 
for computations. An example of a system that creates incomplete 
Gr\accent127obner basis is `Opal'.

\beginexample 
\endexample


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\Section{Attributes and Operations for Groebner Basis}

\>CompletelyReduce( <GB>, <a> ) O

reduces an algebra element $a$ by the Gr\accent127obner basis $GB$. 
The reduced element is returned. If $a$ is already completely 
reduced, the original element $a$ is returned.

\beginexample 
\endexample


\>CompletelyReduceGroebnerBasis( <GB> ) O

modifies a Gr\accent127obner basis $GB$ such that each relation in 
$GB$ is completely reduced.  The `IsCompletelyReducedGroebnerBasis' 
and `IsTipReducedGroebnerBasis' properties are set as a result of 
this operation. The resulting relations will be placed in sorted 
order according to the ordering of $GB$.

\beginexample 
\endexample


\>TipReduce( <GB>, <a> ) O

returns an algebra element $a$ tip reduced by the Gr\accent127obner 
basis $GB$. If $a$ is already tip reduced, then the original 
$a$ is returned.

\beginexample 
\endexample


\>TipReduceGroebnerBasis( <GB> ) O

creates an equivalent Gr\accent127obner basis to $GB$ such that each relation 
generating $GB$ is tip reduced. If $GB$ is already tip reduced, this 
function returns the original object $GB$, possibly with the addition 
of the `IsTipReduced`' property set.

\beginexample 
\endexample


\>Iterator( <GB> ) O

creates an iterator that iterates over the relations making up 
the Gr\accent127obner basis. These relations are iterated over in 
ascending order with respect to the ordering for the
family the elements are contained in.

\beginexample 
\endexample


\>Enumerator( <GB> ) O

creates an enumerator that enumerates the
relations making up the Gr\accent127obner basis. These relations should are
enumerated in ascending order with respect to the ordering for the
family the elements are contained in.

\beginexample 
\endexample


\>Nontips( <GB> ) O

returns a list of nontip elements for a
Gr\accent127obner basis. In order to compute the nontip elements, the
Gr\accent127obner  basis must be complete and tip reduced, and there must be a 
finite number of nontips. If there are an infinite number of nontips,
the operation returns `fail'.

\beginexample 
\endexample


\>AdmitsFinitelyManyNontips( <GB> ) O

returns `true' if the
Gr\accent127obner basis admits only finitely many nontips and `false' 
otherwise. This operation only applies to complete Gr\accent127obner basis.

\beginexample 
\endexample


\>NontipSize( <GB> ) O

returns the number of nontips admitted by 
the Gr\accent127obner basis $GB$. This operation is available only to
complete  basis

\beginexample 
\endexample


\>IsPrefixOfTipInTipIdeal( <GB>, <R> ) O

checks The tip of a relation $R$ to see if it is in the tip ideal 
generated by the tips of $GB$. This is used mainly for the
construction of right Gr\accent127obner basis, but is made available for
general use in case there are other unforseen applications.

\beginexample 
\endexample


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\Section{Right Groebner Basis}

In this section we support right Gr\accent127obner basis for two-sided ideals
with Gr\accent127obner basis. More general cases may be supported in the future.

\>IsRightGroebnerBasis( <GB> ) P

return `true' if an object is a right Gr\accent127obner basis.

\beginexample 
\endexample


\>RightGroebnerBasisOfIdeal( <I> ) A

stores a right Gr\accent127obner basis of a right ideal, $I$,
if one has been computed.

\>RightGroebnerBasis( <R> ) O

The `RightGroebnerBasis' operation constructs a right Gr\accent127obner basis
for the right ideal, $R$,  which must support a right Gr\accent127obner basis 
theory.  Right now, this requires that R have a complete Gr\accent127obner basis.





